---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filibustr

<!-- badges: start -->
[![R-CMD-check](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/feinleib/filibustr/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/feinleib/filibustr/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The `filibustr` package provides data utilities for research on the U.S. Congress.
This package provides a uniform interface for accessing data on members and votes.

## Installation

You can install the development version of filibustr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("feinleib/filibustr")
```

## Functions

### Voteview

There are three functions that retrieve data from [Voteview](https://voteview.com/data):

* `get_voteview_members()`: data on members (Presidents, Senators, and Representatives).
* `get_voteview_parties()`: data on parties (size and ideology)
* `get_voteview_rollcall_votes()`: results of recorded votes (overall results, not positions of individual members)
* `get_voteview_member_votes()`: individual members' votes on recorded votes

These functions share a common interface. Here are their arguments:

* `local`: `r lifecycle::badge('experimental')` Whether to read the data from a local file, as opposed to the Voteview website. Default is `TRUE`. If the local file does not exist, will fall back to reading from Voteview.
* `local_dir`: `r lifecycle::badge('experimental')` The directory containing the local file. Defaults to the working directory.

**Note:** Especially when working with large datasets, reading data from Voteview can take a long time. If you are repeatedly loading the same static dataset (i.e., not including information from the current Congress), it may be useful to download the dataset as a CSV from Voteview so you can read that local file instead of having to use the web interface.

* `chamber`: Which chamber to get data for. Options are:
  * `"all"`, `"congress"`: Both House and Senate data (the default).
  * `"house"`, `"h"`, `"hr"`: House data only.
  * `"senate"`, `"s"`, `"sen"`: Senate data only.
These options are case-insensitive. If you explicitly pass a different value, it will default to "all" with a warning.

**Note:** For `get_voteview_members()` and `get_voteview_parties()`, presidents are included in all datasets. Therefore, reading *both* `"house"` and `"senate"` data will duplicate data on the presidents. The recommended way to get all data is to use the default argument, `"all"`.

* `congress`: A whole number (to get data for a single Congress), or a numeric vector (to get data for a set of congresses). Optional; will retrieve data for all Congresses by default. If specified, Congress numbers cannot be greater than the `current_congress()` (i.e., you cannot try to get future data).

Here is the table returned by `get_voteview_members()`.

```{r Voteview example}
library(filibustr)

get_voteview_members()
```

### Harbridge-Yong (LHY) et al. (2023)

The function `get_lhy_data()` retrives replication data for [Harbridge-Yong et al. (2023)](https://doi.org/10.1086/723805), aka "LHY et al."

`get_lhy_data()` takes the following arguments:

* `local`, `local_dir`: Same as the Voteview functions.
* `chamber`: Which chamber to get data for. See the **Voteview** section above for more info on this argument.

**Note:** Unlike the Voteview functions, there is no "all" option. The House and Senate data do not have the same number of variables, or the same variable names, so it is not trivial to join the two tables.

You must specify either House or Senate data, since there is no "default" option.


Here are the tables returned by `get_lhy_data()`:

```{r LHY example}
library(filibustr)

get_lhy_data("house")
get_lhy_data("senate")
```

### Senate.gov

The following functions retrieve data tables from [Senate.gov](https://www.senate.gov).

* `get_senate_sessions()`: The start and end dates of each legislative session of the Senate. ([table link]((https://www.senate.gov/legislative/DatesofSessionsofCongress.htm)))
* `get_senate_cloture_votes()`: Senate action on cloture motions and cloture votes. ([table link](https://www.senate.gov/legislative/cloture/clotureCounts.htm))

These functions take no arguments, and they always return the full data table from the Senate website.

### Small utilities

This package also provides some smaller utility functions for working with congressional data.

* `year_of_congress()` returns the starting year for a given Congress.
* `congress_in_year()` returns the Congress number for a given year.
* `current_congress()` returns the number of the current Congress, which is currently `r current_congress()`. `current_congress()` is equivalent to `congress_in_year(Sys.Date())`.

## Data sources

This package uses data from the following websites and research:

* Harbridge-Yong, L., Volden, C., & Wiseman, A. E. (2023). The Bipartisan Path to Effective Lawmaking. *The Journal of Politics*, *85*(3), 1048–1063. <https://doi.org/10.1086/723805>
* Lewis, Jeffrey B., Keith Poole, Howard Rosenthal, Adam Boche, Aaron Rudkin, and Luke Sonnet (2023). *Voteview: Congressional Roll-Call Votes Database.* <https://voteview.com/>
* U.S. Senate. <https://www.senate.gov/>

