---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filibustr

<!-- badges: start -->
[![R-CMD-check](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/feinleib/filibustr/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/feinleib/filibustr/actions/workflows/test-coverage.yaml)
[![Codecov test coverage](https://codecov.io/gh/feinleib/filibustr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/feinleib/filibustr?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/filibustr)](https://CRAN.R-project.org/package=filibustr)
<!-- badges: end -->

The [`filibustr`](https://feinleib.github.io/filibustr/) package provides data utilities for research on the U.S. Congress.
This package provides a uniform interface for accessing data from sources such as Voteview, the Legislative Effectiveness Scores, and more. Accessing your data using these functions removes many of the manual steps involved with importing data. This has two primary benefits: 

* **Speeding up your workflow** and enabling you to quickly experiment with a variety of data choices.
* Ensuring you always have the **most up-to-date data**.

`filibustr` is inspired by the [`baseballr`](https://github.com/BillPetti/baseballr) package, which provides similar conveniences for baseball analytics data.

## Installation

You can install the stable version of filibustr from CRAN with:

``` r
install.packages("filibustr")
```

You can install the development version of filibustr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("feinleib/filibustr")
```

## Functions

### Voteview

There are four functions that retrieve data from [Voteview](https://voteview.com/data):

* `get_voteview_members()`: data on members (Presidents, Senators, and Representatives).
* `get_voteview_parties()`: data on parties (size and ideology)
* `get_voteview_rollcall_votes()`: results of recorded votes (overall results, not positions of individual members)
* `get_voteview_member_votes()`: individual members' votes on recorded votes

These functions share a common interface and arguments.

For demonstration, here is the table returned by `get_voteview_parties()`.

```{r Voteview example}
library(filibustr)

get_voteview_parties()
```

**Note:** Especially when working with large datasets, reading data from Voteview can take a long time. Here are two strategies to speed up your data import:

* If you are repeatedly loading the same static dataset (i.e., not including information from the current Congress), it may be useful to download the dataset as a CSV/DTA file so you can read that local file using `local_path` instead of having to download data from online.
* You may use `mirai` to download Voteview data in parallel. See `vignette("parallel-downloads", package = "filibustr")` for more info on parallel data downloads.

### Legislative Effectiveness Scores

The function `get_les()` retrieves Legislative Effectiveness Scores data from the [Center for Effective Lawmaking](https://thelawmakers.org).

Here is an example table returned by `get_les()`.

```{r LES example}
get_les(chamber = "senate")
```

There are non-trivial differences between the House and Senate datasets, so take care when joining House and Senate data.

### Harbridge-Yong, Volden, and Wiseman (2023)

The function `get_hvw_data()` retrives replication data for [Harbridge-Yong, Volden, and Wiseman (2023)](https://doi.org/10.1086/723805).

Here are the tables returned by `get_hvw_data()`:

```{r HVW example}
get_hvw_data("house")
get_hvw_data("senate")
```

The House and Senate data do not have the same number of variables, or the same variable names, so it is not trivial to join the two tables.

### Senate.gov

The following functions retrieve data tables from [Senate.gov](https://www.senate.gov).

* `get_senate_sessions()`: The start and end dates of each legislative session of the Senate. ([table link](https://www.senate.gov/legislative/DatesofSessionsofCongress.htm))
* `get_senate_cloture_votes()`: Senate actions on cloture motions and cloture votes. ([table link](https://www.senate.gov/legislative/cloture/clotureCounts.htm))

These functions take no arguments, and they always return the full data table from the Senate website.

### Small utilities

This package also provides some smaller utility functions for working with congressional data.

* 3 functions dealing with years and Congress numbers:
  * `year_of_congress()` returns the starting year for a given Congress.
  * `congress_in_year()` returns the Congress number for a given year.
  * `current_congress()` returns the number of the current Congress, which is currently `r current_congress()`. `current_congress()` is equivalent to `congress_in_year(Sys.Date())`.
* `get_voteview_cast_codes()` returns a key to the `cast_code` column in `get_voteview_member_votes()`.
* `read_html_table()` is a general-use function for reading HTML tables from online. It's a nice shortcut for a common `rvest` workflow that otherwise takes 3 functions. (It's what powers the Senate.gov functions!)

## Feedback and contributions

If you notice any bugs, or have suggestions for new features, please submit an issue on the [Issues page](https://github.com/feinleib/filibustr/issues) of this package's GitHub repository!

## Data sources

This package uses data from the following websites and research:

* Harbridge-Yong, L., Volden, C., & Wiseman, A. E. (2023). The Bipartisan Path to Effective Lawmaking. *The Journal of Politics*, *85*(3), 1048–1063. <https://doi.org/10.1086/723805>
* Lewis, Jeffrey B., Keith Poole, Howard Rosenthal, Adam Boche, Aaron Rudkin, and Luke Sonnet (`r format(Sys.Date(), "%Y")`). *Voteview: Congressional Roll-Call Votes Database.* <https://voteview.com/>
* U.S. Senate. <https://www.senate.gov/>
* Volden, C., & Wiseman, A. E. (2025). *Legislative Effectiveness Scores* [dataset]. Center for Effective Lawmaking. <https://thelawmakers.org/>
