---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filibustr

<!-- badges: start -->
<!-- badges: end -->

The `filibustr` package provides data utilities for research on the U.S. Congress.
This package provides a uniform interface for accessing data on members and votes.

## Installation

You can install the development version of filibustr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("feinleib/filibustr")
```

## Usage

The `get_voteview_members()` function downloads data from [Voteview](https://voteview.com/data) on Presidents, Senators, and Representatives.

```{r example}
library(filibustr)

get_voteview_members()
```

## Data sources

This package uses data from the following websites and research:
* Lewis, Jeffrey B., Keith Poole, Howard Rosenthal, Adam Boche, Aaron Rudkin, and Luke Sonnet (2023). *Voteview: Congressional Roll-Call Votes Database.* <https://voteview.com/>
* Harbridge-Yong, L., Volden, C., & Wiseman, A. E. (2023). The Bipartisan Path to Effective Lawmaking. *The Journal of Politics*, *85*(3), 1048–1063. <https://doi.org/10.1086/723805>


