---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filibustr

<!-- badges: start -->
[![R-CMD-check](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/feinleib/filibustr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `filibustr` package provides data utilities for research on the U.S. Congress.
This package provides a uniform interface for accessing data on members and votes.

## Installation

You can install the development version of filibustr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("feinleib/filibustr")
```

## Functions

### Voteview

There are three functions that download data from [Voteview](https://voteview.com/data):

* `get_voteview_members()`: data on members (Presidents, Senators, and Representatives).
* `get_voteview_parties()`: data on parties (size and ideology)
* `get_voteview_rollcall_votes()`: results of recorded votes (overall results, not positions of individual members)

These functions share a common interface. Here are their arguments:

* `local`: Whether to read the data from a local file, as opposed to the Voteview website. Default is `TRUE`. If the local file does not exist, will fall back to reading from Voteview.
* `local_dir`: The directory containing the local file. Defaults to the working directory.
* `chamber`: Which chamber to get data for. Options are:
  * `"all"`, `"congress"`: Both House and Senate data (the default).
  * `"house"`, `"h"`, `"hr"`: House data only.
  * `"senate"`, `"s"`, `"sen"`: Senate data only.
These options are case-insensitive. If you explicitly pass a different value, it will default to "all" with a warning.

**Note:** for `get_voteview_members()` and `get_voteview_parties()`, presidents are included in all datasets. Therefore, reading *both* `"house"` and `"senate"` data will duplicate data on the presidents. The recommended way to get all data is to use the default argument, `"all"`.

* `congress`: A whole number (to get data for a single Congress), or a numeric vector (to get data for a set of congresses). Optional; will retrieve data for all Congresses by default. If specified, Congress numbers cannot be greater than the `current_congress()` (i.e., you cannot try to get future data).

Here is the table returned by `get_voteview_members()`.

```{r Voteview example}
library(filibustr)

get_voteview_members()
```

### Senate.gov

The function `get_senate_sessions()` retrieves the start and end dates of each legislative session of the Senate.

This function takes no arguments, as it always returns the full [table from the Senate website](https://www.senate.gov/legislative/DatesofSessionsofCongress.htm).

### Small utilities

This package also provides some smaller utility functions for working with congressional data.

* `current_congress()` returns the number of the current Congress, which is currently `r current_congress()`.

## Data sources

This package uses data from the following websites and research:

* Harbridge-Yong, L., Volden, C., & Wiseman, A. E. (2023). The Bipartisan Path to Effective Lawmaking. *The Journal of Politics*, *85*(3), 1048–1063. <https://doi.org/10.1086/723805>
* Lewis, Jeffrey B., Keith Poole, Howard Rosenthal, Adam Boche, Aaron Rudkin, and Luke Sonnet (2023). *Voteview: Congressional Roll-Call Votes Database.* <https://voteview.com/>
* U.S. Senate. <https://www.senate.gov/>

